The core transcriptome of mammalian placentas and the divergence of expression with placental shape

### 2017-02-13 manuscript/manuscript.Rnw
discuss TSI in manuscript; move enrichment table

The
\Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
genes comprising the core placenta transcriptome are shown in
\cref{fig:core_placenta_transcriptome} ordered by their median
transcription level in \ac{FPKM}. We considered a gene to be core to
the placenta transcriptome if a single-copy ortholog of that gene was
expressed with $\text{FPKM} \ge 10$ in all eutherian species studied
(excluding the metatherian, \opossum) and was not a human housekeeping
gene according to
\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}. Genes which
meet this criteria but are housekeeping genes are depicted in
Supplemental \cref{fig:core_placenta_transcriptome_housekeeping}.
\Cref{fig:core_placenta_transcriptome_shape} shows the number of genes
which would be included in the core placenta set if the \ac{FPKM}
cutoff was set differently; for example, choosing an \ac{FPKM} cutoff
of 1 yields %
\Sexpr{cpt.distribution.shape[`Expression (FPKM)`>=1 & `Min. Exp. in`
  == '14 of 14 Species',max(num_genes)]} %
genes in the core transcriptome. %
In contrast to the shape of the number of genes included in the core
placenta transcriptome, the set of \human housekeeping genes expressed
all placentas is flatter (Supplemental
\cref{fig:core_housekeeping_transcriptome_shape}), and includes more
of the total fraction of genes. However, there is still a large
difference between the set of housekeeping genes which are expressed
in \human and those which are expressed in all placentas.

>>>>

The
\Sexpr{core.placenta.transcriptome.long[housekeeping==FALSE,length(unique(human_name))]}
genes comprising the core placenta transcriptome are shown in
\cref{fig:core_placenta_transcriptome} ordered by their median
transcription level in \ac{FPKM}. We considered a gene to be core to
the placenta transcriptome if a single-copy ortholog of that gene was
expressed with $\text{FPKM} \ge 10$ in all eutherian species studied
(excluding the metatherian, \opossum) and was not a human housekeeping
gene according to
\citet{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}. Genes which
meet this criteria but are housekeeping genes are depicted in
Supplemental \cref{fig:core_placenta_transcriptome_housekeeping}.
\Cref{fig:core_placenta_transcriptome_shape} shows the number of genes
which would be included in the core placenta set if the \ac{FPKM}
cutoff was set differently; for example, choosing an \ac{FPKM} cutoff
of 1 yields %
\Sexpr{cpt.distribution.shape[`Expression (FPKM)`>=1 & `Min. Exp. in`
  == '14 of 14 Species',max(num_genes)]} %
genes in the core transcriptome. %
In contrast to the shape of the number of genes included in the core
placenta transcriptome, the set of \human housekeeping genes expressed
all placentas is flatter (Supplemental
\cref{fig:core_housekeeping_transcriptome_shape}), and includes more
of the total fraction of genes. An alternative approach to defining
housekeeping genes is to use the tissue specificity
index~\cite{Yanai.ea2005:Genome-widemidrangetranscriptionprofiles}.
Supplemental \cref{tab:tissue_specificity_index_of_plaenta_core_genes}
lists the tissue specificity index for all housekeeping and
non-housekeeping genes; housekeeping genes expressed in all placentas
have a lesser median tissue specificity index than non-housekeeping
genes (Supplemental
\cref{fig:tissue_specificity_index_with_housekeeping}).
